%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[letterpaper]{tufte-handout}

%\geometry{showframe} % display margins for debugging page layout

\usepackage{fontspec}
\def\mainfont{Linux Libertine O}
\setmainfont[Ligatures={Common,TeX}, Contextuals={NoAlternate}, BoldFont={* Bold}, ItalicFont={* Italic}, Numbers={OldStyle}]{\mainfont}
\setsansfont[Scale=MatchLowercase, Numbers={OldStyle}]{Linux Biolinum O} 
\setmonofont{Linux Libertine O}
\usepackage{microtype}

\usepackage{graphicx} % allow embedded images
  \setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
  \graphicspath{{img/}} % set of paths to search for images
\usepackage{amsmath}  % extended mathematics
\usepackage{booktabs} % book-quality tables
\usepackage{units}    % non-stacked fractions and better unit spacing
\usepackage{multicol} % multiple column layout facilities
%\usepackage{fancyvrb} % extended verbatim environments
%  \fvset{fontsize=\normalsize}% default font size for fancy-verbatim environments

\usepackage{enumitem}
\usepackage{mhchem}

\makeatletter
% Paragraph indentation and separation for normal text
\renewcommand{\@tufte@reset@par}{%
  \setlength{\RaggedRightParindent}{1.0pc}%
  \setlength{\JustifyingParindent}{1.0pc}%
  \setlength{\parindent}{1pc}%
  \setlength{\parskip}{0pt}%
}
\@tufte@reset@par

% Paragraph indentation and separation for marginal text
\renewcommand{\@tufte@margin@par}{%
  \setlength{\RaggedRightParindent}{0pt}%
  \setlength{\JustifyingParindent}{0.5pc}%
  \setlength{\parindent}{0.5pc}%
  \setlength{\parskip}{0pt}%
}
\makeatother

% Set up the spacing using fontspec features
   \renewcommand\allcapsspacing[1]{{\addfontfeatures{LetterSpace=15}#1}}
   \renewcommand\smallcapsspacing[1]{{\addfontfeatures{LetterSpace=10}#1}}

\title{{\scshape bi} 163 Study Guide 10}

\date{} % without \date command, current date is supplied

\begin{document}

\maketitle	% this prints the handout title, author, and date

%\printclassoptions
\section*{Transitional forms 1}

We\marginnote{\textbf{Read:} 531, 536--537, 542--544, 676--672, 728--731, 738--739, 741--742.} covered evidence for transitional forms that represent major macroevolutionary events.

\section*{Vocabulary}

\vspace{-1\baselineskip}
\begin{multicols}{2}
Ediacaran Period\\
Cambrian Period\\
Cambrian radiation\\
trace fossils \\
adaptive radiation \\
tetrapods \\
\textit{Eusthenopteron} \\
\textit{Ichthyostega} \\
\textit{Tiktaalik} \\
articular-quadrate jaw joint \\
dentary-squamosal jaw joint\\
\textit{Archaeopteryx} \\
\end{multicols}

\section*{Concepts}

You should \emph{write} clear and concise answers to each question in the Concepts section.  Remember to ``think horizontally'' and to ``connect the dots.'' \vspace*{\baselineskip}

Hold on tight. This is a long one!

\begin{enumerate}

	\item When did the Ediacaran period occur?

	\item When did the Ediacaran fauna live, based on fossil evidence?\marginnote{The Ediacaran fauna was unique to the that period, based on unique morphological features and body plans. None of the Ediacaran fauna lived past the early Cambrian Period so none of the organisms alive today are descendants of the that fauna. The earliest ancestors of life today do trace back to the Ediacaran Period—and even earlier—but those ancestors are \emph{not} considered part of the Ediacaran fauna.}

	\item What did the Cambrian Period begin?
	
	\item What is the Cambrian radiation?  Why is this a better term than the often heard Cambrian explosion?
	
	\item What are trace fossils? How old are the oldest trace fossils ever found on Earth? What is the difference between the horizontal and vertical trace fossils found at the end of the Ediacaran and the beginning of the Cambrian periods?
	
	\item What is an adaptive radiation? \marginnote{\url{http://www.nature.com/news/what-sparked-the-cambrian-explosion-1.19379}} How does this relate to the Cambrian radiation. How could an adaptive radiation be related to changing environmental conditions (like additional \ce{O2} in the atmosphere) or a \emph{prossible} mass extinction  at the end of the Ediacaran Period? How could  these types of events set the stage for an adaptive radiation early in the Cambrian Period?
	
	\item What is a transitional fossil or transitional form?

	\item Describe \marginnote{Many other transitional fossils have also been found between \textit{Eusthenopteron} and \textit{Ichthyostega} besides \textit{Tiktaalik} but this one showed many important transitional features.} the evidence for transitional forms between \textit{Eusthenopteron} and \textit{Ichthyostega}. What was name of the important transitional fossil between \textit{Eusthenopteron} and \textit{Ichthyostega}.

	
	\item When did \textit{Tiktaalik} live? How does this support it being a transitional form between \textit{Eusthenopteron} and \textit{Ichthyostega}?
	
	\item What \marginnote{\textit{Eusthenopteron} was a lobe-finned fish. \textit{Ichthyostega} was a land-dwelling tetrapod.} are some of the morphological features of \textit{Tiktaalik} lthat support it being a transitional form between \textit{Eusthenopteron} and \textit{Ichthyostega}?
	
	\item What are some of the traits found in reptile-like mammals that support the hypothesis that they are transitional fossils between reptilian ancestors and mammalian descendants?
	
	\item Explain the difference between the articular-quadrate and dentary-squamosal jaw joints. Which type of jaw joint is found in reptiles? In mammals?
	
	\item What became of the articular and quadrate bones that formed the reptilian jaw joint.
	
	\item What morphological features did \textit{Archaeopteryx} have that support it being a transitional fossil between dinosaurs and birds? Could \textit{Archaeopteryx} fly? Why or why not? 
	
	\item Give some reasons for the possible evolution of feathers. 
	
	\item What evidence suggests that \textit{Tyrannosaurs rex} \marginnote{\url{http://news.nationalgeographic.com/news/2012/04/120404-yutyrannus-feathers-dinosaur-science-nature-biggest/}}might have had feathers? Does this make sense in terms of monophyletic groups and common ancestry? Explain why or why not.
	
	\item What genetic evidence suggests that birds are evolved from dinosaurs?
	
	\item Which came first, the chicken or the egg? Explain it in biological terms of common ancestry.
	
	\item Are dinosaurs extinct? Justify your answer in terms of monophyletic groups and common ancestry.
	
\end{enumerate}

\end{document}