%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt]{exam}
\usepackage{graphicx}
	\graphicspath{{/Users/goby/Pictures/teach/163/lab/}
	{img/}} % set of paths to search for images

\usepackage{geometry}
\geometry{letterpaper, left=1.5in, bottom=1in}                   
%\geometry{landscape}                % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{amssymb, amsmath}
\usepackage{mathtools}
	\everymath{\displaystyle}

\usepackage{fontspec}
\setmainfont[Ligatures={TeX}, BoldFont={* Bold}, ItalicFont={* Italic}, BoldItalicFont={* BoldItalic}, Numbers={OldStyle}]{Linux Libertine O}
\setsansfont[Scale=MatchLowercase,Ligatures=TeX]{Linux Biolinum O}
\setmonofont[Scale=MatchLowercase]{Inconsolatazi4}
\usepackage{microtype}


% To define fonts for particular uses within a document. For example, 
% This sets the Libertine font to use tabular number format for tables.
 %\newfontfamily{\tablenumbers}[Numbers={Monospaced}]{Linux Libertine O}
% \newfontfamily{\libertinedisplay}{Linux Libertine Display O}

\usepackage{booktabs}
\usepackage{multicol}
\usepackage[normalem]{ulem}

\usepackage{longtable}
%\usepackage{siunitx}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}

\usepackage{enumitem}
\setlist{leftmargin=*}
\setlist[1]{labelindent=\parindent}

\usepackage{hyperref}
%\usepackage{placeins} %PRovides \FloatBarrier to flush all floats before a certain point.
\usepackage{hanging}

\usepackage[sc]{titlesec}

%% Commands for Exam class
\renewcommand{\solutiontitle}{\noindent}
\unframedsolutions
\SolutionEmphasis{\bfseries}

\renewcommand{\questionshook}{%
	\setlength{\leftmargin}{-\leftskip}%
}

%Change \half command from 1/2 to .5
\renewcommand*\half{.5}

\pagestyle{headandfoot}
\firstpageheader{\textsc{bi}\,063 Evolution and Ecology}{}{\ifprintanswers\textbf{KEY}\else Name: \enspace \makebox[2.5in]{\hrulefill}\fi}
\runningheader{}{}{\footnotesize{pg. \thepage}}
\footer{}{}{}
\runningheadrule

\newcommand*\AnswerBox[2]{%
    \parbox[t][#1]{0.92\textwidth}{%
    \begin{solution}#2\end{solution}}
%    \vspace*{\stretch{1}}
}

\newenvironment{AnswerPage}[1]
    {\begin{minipage}[t][#1]{0.92\textwidth}%
    \begin{solution}}
    {\end{solution}\end{minipage}
    \vspace*{\stretch{1}}}

\newlength{\basespace}
\setlength{\basespace}{5\baselineskip}


%\usepackage{mdframed}
%\mdfsetup{%
%	innerleftmargin=0pt,%
%	innerrightmargin=0pt,
%	innertopmargin=0pt,
%	innerbottommargin=0pt,
%	hidealllines=true
%}%end mdfsetup

%
%\makeatletter
%\def\SetTotalwidth{\advance\linewidth by \@totalleftmargin
%\@totalleftmargin=0pt}
%\makeatother


%\printanswers


\begin{document}

\subsection*{Journal 2: evaluating all of the evidence (60 points)}

\emph{Read this entire handout throughly. You are responsible for meeting all requirements of this assignment, as detailed below.}

This is your second journal assignment, which is a formal writing assignment. 
This journal entry will require
some writing, as you need to explain the effect of the transitional forms and genetic evidence on your hypothesis. It isn't really that hard,
though. Look at your phylogenetic hypothesis to see which organisms you
predicted are related (or not). Your answer to this assignment should
take \emph{at least} two typed pages (double space, please!), and
probably a little more, depending on your hypothesis.

You have now seen evidence for transitional forms in the
fossil record. You have also seen evidence of relationships through 
analyses of \textsc{dna} sequences. Does this evidence support or falsify your hypothesis? If it
supports your hypothesis, tell how. If it falsifies your hypothesis,
tell how, and tell how you would have to revise your hypothesis to fit
the evidence. Remember to include statements that clearly identify the
predictions made by your hypothesis and whether the evidence from
transitional forms agrees or disagrees with the predictions, and then
tell whether your hypothesis was supported or falsified.

This is the culmination of all of the evidence provided in lab to test your hypothesis.
The evidence is listed below. Also listed is the component point value of each item for this assignment.

\begin{enumerate}

	\item Fish / Amphibian Transition (5 points).

	\item Reptile / Mammal transition (5 points).
	
	\item Reptile / Bird Transition (5 points).

	\item Whale / Even-Toed Hooved Mammal Transition (5 points).

	\item Hominid Transitions (humans and great apes) (5 points)

	\item Geological Time Scale (5 points): 
	
	A geological time scale is available on the \emph{lecture}
	Moodle page. You \emph{must} use this time scale to evaluate your 
	last hypothesis and make your final hypothesis. 
	
	The time scale shows when different groups of 
	organisms (mammals, insects, single-celled eukaryotes, etc.)
	first appeared in the fossil record. Did your predictions about the time of
	appearance for each group of organisms (mammals, reptiles, birds, etc.)
	agree or disagree with the fossil record and the geological time scale? 
	Carefully consider the order of appearance of the different organismal 
	groups. This will be critical to getting a correct hypothesis. 
  
	\emph{Important:} The fossil record also shows evidence of transitional 
	forms from amphibians to reptiles. We did not cover this evidence in class 
	because to understand the differences requires more detail than we can cover in this class. 

	\item Genetic analysis and phylogenetic trees (5 points).
	
\end{enumerate}

Your final phylogenetic tree must also be consistent with the 
anatomical, embryological, and larval homologies but you do not have to discuss
that evidence in this journal assignment. You must discuss the new evidence listed above.

\subsubsection*{Assignment}

Before you begin, review all of your assignments. Compare the evidence from
the lab exercises to the predictions made by your phylogenetic trees. Does the 
evidence support your hypothesis? Falsify your hypothesis? Give no conclusive 
evidence? Does the evidence require you to change your hypothesis in any way? Refer to the Scientific Method
Overview for how to use evidence to test hypotheses.

Some parts of your hypothesis may be supported and
some parts may be falsified by all of the evidence presented during in lab.
You must discuss both supported and unsupported parts of your
hypothesis. For supported parts, you must discuss which parts of your
hypothesis are supported and how the evidence supports your hypothesis.
For falsified parts of your hypothesis, you must explain which parts are
not supported, how the evidence falsifies your hypothesis, and then
explain how your hypothesis must be revised to fit the evidence examined
so far. You have evaluated evidence for all organisms so you must include all organisms in your discussion.

\emph{Clarity of your written thoughts is critical for your journal entry. Writing that is not clear is evidence of
 thinking that is not clear and will be evaluated accordingly.}

\subsubsection*{Grading of Journal 2}

\begin{enumerate}

\item Evaluation of the evidence (35 points)

Each type of evidence that you evaluated in class has been given a point
value (see the evidence above). The points will be assigned based on
inclusion and thorough discussion of the evidence, and how the evidence
affected your hypothesis. Did the evidence support the predictions made
by your hypothesis? Did the evidence falsify your hypothesis? If the
evidence falsified your hypothesis, then you must state how you revised
your hypothesis to agree with the evidence. Clarity counts towards the
total points in each category.

\item Spelling, grammar, and mechanics (5 points)

Spelling, grammar and
mechanics (sentence structure, etc) are important in any writing but especially informal writing. Use spell check. Proof your assignment carefully before you turn it in. You are entitled to \emph{one free
mistake.} After that, each mistake is a 1 point deduction.

\emph{Remember}: This course is part of the University Studies
program. The University Studies program objectives that are especially
relevant to this course are 1) the demonstration of your ability for
critical thinking, reasoning and analyzing, and 2) the demonstration of
effective written communication. Keep this in mind as you write. Points will be deducted for signs of
weak reasoning and analysis and poorly written communication.

\item New phylogenetic tree (20 points)

You must also submit a revised phylogenetic tree with this assignment. 
Your new phylogenetic tree must be consistent with every piece of
evidence evaluated in lab. It must also be consist with what you write
for your journal entry assignment above. You will lose 3
points for each missed homology or transitional form (or time of transition). 
Use the phylogenetic trees you made using \textsc{dna} sequence data as a guide.
You will also lose 3 points for each missing organism, a missing or incomplete time scale, 
or any other common errors. \emph{See the first phylogenetic tree exercise for things you should not do.}

\end{enumerate}


\subsubsection*{Due Date} 

Your journal entry and new phylogenetic tree are due at the start of lab next week. Late submissions will not be accepted without a valid excuse. If necessary, review the syllabus for lab policy.

Please do not type a cover page. Just type your name, \textsc{bi} 063 and section number at the
top, and then start typing your journal assignment.

\end{document}  