%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt]{exam}


%\printanswers

\usepackage{geometry}
\geometry{letterpaper, left=1.5in, bottom=1in}                   
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent

\usepackage{fontspec}
\setmainfont[Ligatures={TeX}, BoldFont={* Bold}, ItalicFont={* Italic}, BoldItalicFont={* BoldItalic}, Numbers={OldStyle}]{Linux Libertine O}
\setsansfont[Scale=MatchLowercase,Ligatures=TeX]{Linux Biolinum O}
\setmonofont[Scale=MatchLowercase]{Inconsolatazi4}
\newfontfamily{\tablenumbers}[Numbers={Monospaced,Lining}]{Linux Libertine O}
\usepackage{microtype}

\usepackage{enumitem}
\setlist{leftmargin=*}
\setlist[1]{labelindent=\parindent}
\setlist[enumerate]{label=\textsc{\alph*}.}
\setlist[itemize]{label=\color{gray}\textbullet}

\usepackage[sc]{titlesec}


\pagestyle{headandfoot}
\firstpageheader{\textsc{bi}\,163 Evolution and Ecology}{}{\ifprintanswers\textbf{KEY}\else Name: \enspace \makebox[2.5in]{\hrulefill}\fi}
\runningheader{}{}{\footnotesize{pg. \thepage}}
\footer{}{}{}
\runningheadrule


\begin{document}

\subsection*{Search for the missing sea otters, part 2 (16 points)}

You have learned that sea otter populations have crashed at some Aleutian Islands in much less than a decade. The sea otter population could be declining for a number of reasons. Follow these instructions carefully.


\subsubsection*{Assignment}

\begin{enumerate}
\item Type the four rates that determine whether the population growth rate is increasing (positive $r$) or decreasing (negative $r$). A list of the four rates will suffice. Review your notes and slides for Lecture 13 if you do not remember the four rates. (4 points)

\item Write one hypothesis for each of the four rates that might explain sea otter population decline.You must write one hypothesis for each rate, for a total of four hypothesis. Remember that a hypothesis is a direct statement that tentatively explains an observation. Each hypothesis must be a complete sentence. (3 points per hypothesis.)

\item Upload your typed assignment to the Missing Otters Part 2 drop box by the due date. You must upload your assignment before you can continue with Part 3 of this case study.

\end{enumerate}

\textbf{Your hypotheses must be as specific as possible.} For example, don’t just say “Fewer individuals are giving birth.” What do you think could be specific causes of decreased birth rates?

\end{document}  