%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt, hidelinks, addpoints]{exam}
\usepackage{graphicx}
	\graphicspath{{/Users/goby/Pictures/teach/163/lab/}
	{img/}} % set of paths to search for images

\usepackage{geometry}
\geometry{letterpaper, left=1.5in, bottom=1in}                   
%\geometry{landscape}                % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{amssymb, amsmath}
\usepackage{mathtools}
	\everymath{\displaystyle}

\usepackage{fontspec}
\setmainfont[Ligatures={TeX}, BoldFont={* Bold}, ItalicFont={* Italic}, BoldItalicFont={* BoldItalic}, Numbers={OldStyle}]{Linux Libertine O}
\setsansfont[Scale=MatchLowercase,Ligatures=TeX, Numbers=OldStyle]{Linux Biolinum O}
%\setmonofont[Scale=MatchLowercase]{Inconsolatazi4}
\usepackage{microtype}


% To define fonts for particular uses within a document. For example, 
% This sets the Libertine font to use tabular number format for tables.
 %\newfontfamily{\tablenumbers}[Numbers={Monospaced}]{Linux Libertine O}
% \newfontfamily{\libertinedisplay}{Linux Libertine Display O}

\usepackage{booktabs}
\usepackage{multicol}
\usepackage[normalem]{ulem}

\usepackage{longtable}
%\usepackage{siunitx}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}

\usepackage{enumitem}
\setlist{leftmargin=*}
\setlist[1]{labelindent=\parindent}
\setlist[enumerate]{label=\textsc{\alph*}.}
\setlist[itemize]{label=\color{gray}\textbullet}

\usepackage{hyperref}
%\usepackage{placeins} %PRovides \FloatBarrier to flush all floats before a certain point.
\usepackage{hanging}

\usepackage[sc]{titlesec}

%% Commands for Exam class
\renewcommand{\solutiontitle}{\noindent}
\unframedsolutions
\SolutionEmphasis{\bfseries}

\renewcommand{\questionshook}{%
	\setlength{\leftmargin}{-\leftskip}%
}

\newcommand{\hidepoints}{%
	\pointsinmargin\pointformat{}
}

\newcommand{\showpoints}{%
	\nopointsinmargin\pointformat{(\thepoints)}
}

%Change \half command from 1/2 to .5
\renewcommand*\half{.5}

\pagestyle{headandfoot}
\firstpageheader{\textsc{bi}\,063 Evolution and Ecology}{}{\ifprintanswers\textbf{KEY}\else Name: \enspace \makebox[2.5in]{\hrulefill}\fi}
\runningheader{}{}{\footnotesize{pg. \thepage}}
\footer{}{}{}
\runningheadrule

\newcommand*\AnswerBox[2]{%
    \parbox[t][#1]{0.92\textwidth}{%
    \vspace{-0.5\baselineskip}\begin{solution}\textbf{#2}\end{solution}}
%    \vspace*{\stretch{1}}
}

\newenvironment{AnswerPage}[1]
    {\begin{minipage}[t][#1]{0.92\textwidth}%
    \begin{solution}}
    {\end{solution}\end{minipage}
    \vspace*{\stretch{1}}}

\newlength{\basespace}
\setlength{\basespace}{5\baselineskip}



%\printanswers

\hidepoints

\begin{document}


\subsection*{Evaluating a scientific paper (\numpoints~points)}

\begin{questions}

\question[1]
List the author(s). Name the institution that each author is associated with. 
Besides the author(s), who else contributed to this study? (\textsc{Hint:} You have to look elsewhere in the paper to answer the last part of this question.)

\AnswerBox{3\baselineskip}{%
Diane Wood and Allan Bornstein, both from \textsc{semo}.
}

\question[1]
List three other studies upon which this study was based. Why is it important to look at previous studies? How does this study contribute to the scientific literature?

\AnswerBox{1.75\basespace}{%
Three citations. Previous studies show what is already known. This study adds information on the biology of \textit{Obolaria} and shows that plants that have to grow through lots of leaf litter may affect their ability to produce flowers.
}

\question[1]
What organism was studied? What kind of organism is it?

\AnswerBox{2\baselineskip}{A plant, \textit{Obolaria virginica.}}


\question[1]
What question(s) does the study address? Is a hypothesis
clearly stated? What is the prediction(s) made by the hypothesis?

\AnswerBox{1.75\basespace}{What is the population size of \textit{Obolaria?} What type of dispersal pattern does it have? Does leaf litter affect the number of flowers produced? The hypothesis is that the number of flowers produced by \textit{Obolaria} is related to stem length in the leaf litter. They predicted that plants that have longer stems in the leaf litter will produce fewer flowers. Plants that produce shorter stems in the leaf litter will produce more flowers. }

\question[1]
Where was the study conducted?

\AnswerBox{2\baselineskip}{Trail of Tears State Park, Cape Girardeau County, MO}

\question[1]
Was the study observational or experimental? What type of data were collected? 

\AnswerBox{\basespace}{This was an observational study because no experimental manipulations were performed. The authors counted the number of plants per plot to estimate population size and dispersion. They measured total stem length and above-litter stem length.}


\question[1]
How were the data analyzed? Do the figures and tables provide
appropriate visualization of the data?

\AnswerBox{1.75\basespace}{The data were graphed and a regression line was plotted. Points above the regression line were placed in group 1 (longer stems in the leaf litter). Points below the line were placed in group 2 (longer stems above the litter). A Mann-Whitney U non-parametric test was used to compare flower production between groups 1 and 2.}

\question[1]
What were the results of the study?

\AnswerBox{1.75\basespace}{Population estimates ranged from 833--7820 individuals per 50 $\times$ 50 plot. All dispersion ratios were much greater than 1, indicating a clumped distribution. Plants with longer within-litter stems (group 1) produced significantly fewer flowers than did plants with longer above-litter stems (group 2).}

\question[1]
What were the main conclusions of the study? Did the study address the hypothesis?

\AnswerBox{\basespace}{\textit{Obolaria} may be more common than realized. Large amounts of leaf litter may negatively affect the ability of \textit{Obolaria} to grow and produce flowers. Their hypothesis was supported.}

\question[1]
After reading the paper, do you think that the title adequately
communicates the paper's content?

\AnswerBox{\baselineskip}{Why yes, I do.}


\end{questions}

\end{document}  